An Imaging Search for Circumstellar Companions of Sirius B

### 2021-08-13 paper/paper.tex
update paper

\documentclass[twocolumn]{aastex631}
\usepackage{hyperref}
\let\tablenum\relax
\usepackage{siunitx}
\usepackage{makecell}

\DeclareSIUnit\parsec{pc}
\DeclareSIUnit\au{AU}
\DeclareSIUnit\jupitermass{M_J}
\DeclareSIUnit\milliarcsecond{mas}
\DeclareSIUnit\solarmass{M_\odot}
\DeclareSIUnit\year{yr}
\sisetup{
    list-units=single,
    separate-uncertainty
}

@@

\begin{abstract}
We present deep imaging of Sirius B, the closest and brightest white dwarf. We use Keck/NIRC2 in Lp-band (\qty{3.776}{\micro\meter}) across 3 epochs in 2020 using the technique of angular differential imaging. We reach sub-Jupiter sensitivities and sub-AU separations, reaching \qty{3.5}{\jupitermass} at \qty{0.25}{\au} down to a sensitivity of \qty{0.6}{\jupitermass} at \qty{1.5}{\au}. The uncertainty in mass sensitivity is below \qty{0.1}{\jupitermass} due to the high-precision of the Sirius B system age. We do not detect any companions around Sirius B, consistent with previous studies.
\end{abstract}

@@

High-contrast imaging (HCI) is a powerful technique for discovering and characterizing exoplanets. Being able to probe the architecture, formation, and atmospheres of planets directly is necessary for advancing substellar companion formation and evolution theory. The process required to image a planet is daunting, however. The typical astrophysical flux ratios (contrast) for a Sun-Jupiter analog in the near-infrared (NIR) are $\sim$$10^{\text{-}8}$, and for a Sun-Earth system are $\sim$$10^{\text{-}10}$ \citep{traub_direct_2010}. Blackbody emission from exoplanets peaks in the infrared and is well into the Rayleigh-Jeans limit of the star, decreasing contrast compared to the visible or ultraviolet. In addition, the close angular separations of planets makes it difficult to detect them over the diffraction pattern of their host and other noise sources.

Typical targets for imaging are nearby young stars; the close proximity means the same angular separation probes a smaller physical separation, allowing for studies closer to solar-system scales. Younger exoplanets are hotter and therefore brighter, reducing the flux ratio between the planet and its host star. Another way to reduce contrast would be to image a faint star with bright companions, such as a young white dwarf (WD) star. When intermediate mass stars (\qtyrange{1}{8}{\solarmass}) eventually exit the main-sequence (MS) they expand up to $\sim$100s of stellar radii during the red giant branch (RGB) phase. This expansion will engulf planets within the star's Roche limit. Then, during the asymptotic giant branch (AGB) phase, mass loss will adiabatically expand any orbits. Finally, once the star becomes a white dwarf it begins cooling, which reduces its luminosity by 3 to 4 orders of magnitude compared to its MS progenitor. The faintness, broad spectral lines, and expanded orbits make white dwarfs exceptionally challenging for the transit photometry and radial velocity methods. However, the reduced contrast due to the faintness of the star, and the expanded separation of planets both benefit imaging.

@@

There is limited knowledge of planetary systems around evolved stars. \citet{burleigh_imaging_2002,veras_post-main-sequence_2016} suggest exoplanets on initially wide ($>$\qty{5}{\au}) orbits around intermediate mass stars will survive expansion during the RGB phase. Close-in planets that escape the Roche limit of the expanding red giant can still be shredded by tidal forces, although the tidal forces on planets that readily escape engulfment are negligible \citep{nordhaus_orbits_2013}. During the AGB phase, stellar mass loss will adiabatically expand the orbit of the planet by a maximum factor of $M_{*,\mathrm{MS}}/M_{*,\mathrm{WD}}$ \citep{jeans_cosmogonic_1924}. In addition to \textit{first-generation} planets, there are potential methods for \textit{second-generation} planets to form after the violent RGB and AGB phases \citep{perets_second_2010}.

The first search for substellar companions around white dwarfs was conducted by \citet{probst_infrared_1983} by searching for infrared (IR) excess in their spectral energy distributions (SED) using broadband photometry. They found no companions around the $\sim$100 white dwarfs they studied. The same method was applied by \citet{zuckerman_excess_1987} who found excess IR emission around white dwarf G29-38. Eventually this excess was determined to be from a dust disk \citep{telesco_observations_1990} and the current interpretation associates the dust with accretion onto the white dwarf, polluting the stellar atmosphere \citep{koester_metals_1997}.

To date, there has been no direct images of an exoplanet around a white dwarf, although a brown dwarf with a mass of \qty{7}{\jupitermass} on a wide \qty{2500}{\au} orbit was imaged around a single white dwarf (WD 0806-661B) using Spitzer \citep{luhman_discovery_2011}. The ``Degenerate Objects around Degenerate Objects'' survey \citep[DODO;][]{hogan_dodo_2009} observed 29 white dwarfs with Gemini/NIRI and VLT/ISAAC, reaching an average upper limit around $\sim$\qty{8}{\jupitermass} beyond \qty{35}{\au}. The young white dwarf GD 50 was observed using the extreme AO instrument SPHERE at the VLT \citep{xu_extreme-ao_2015}, reaching sensitivity limits of \qty{4}{\jupitermass} at \qty{6.2}{\au}.

@@

A particularly fascinating such target is Sirius B, the closest and brightest white dwarf. The Sirius system is the 7th closest to the sun at \qty{2.6}{\parsec}, consisting of Sirius A, a -1.35 magnitude A1Vm star known for being the brightest, and Sirius B, a DA2 white dwarf with a \qty{50}{\year} orbit \citep{bond_sirius_2017,gaia_collaboration_gaia_2018}. As mentioned previously, the proximity and faintness of Sirius B (compared to a MS star) make it compelling for imaging, and additionally it is a young system ($\sim$\qty{225}{\mega\year}), further reducing the predicted contrast necessary to image a planetary companion.

The age of the Sirius system has much better precision than most stars ($\sim$1\% vs. $\sim$10\% or worse), which is paramount for estimating the mass of a companion from photometry. Because Sirius B is relatively young, the age uncertainty exponentially affects planet mass uncertainty due to the rapid cooling of young planets ($<$\qty{1}{\giga\year}) \citep{fortney_giant_2010}. The high system age precision is primarily derived from the well-defined dynamical masses of the binary. With well-known masses, isochrones and stellar evolution codes can be combined to pinpoint the age. The dynamical masses are determined through astrometric studies; the first study of Sirius was performed by \citet{bessel_variations_1844}, who recognized wobbles in the proper motion of Sirius A caused by a ``dark satellite''. This dark satellite was visually confirmed in \citet{bond_companion_1862} as Sirius B. \citet{adams_spectrum_1915} took the first spectral measurements of Sirius B and found it to be similar to a MS early A-type star, despite its faintness, which we now know to be typical of white dwarf spectra.

Initial astrometric perturbations suggested a 50 year orbital period \citep{auwers_orbit_1864} for Sirius B. \citet{van_den_bos_orbit_1960,gatewood_study_1978} were the first to estimate dynamic masses using compilations of photographic plates from Lick and Yerkes observatories. \citet{bond_sirius_2017} greatly refined the orbital solution using a compilation of historical data and Hubble Space Telescope (HST) data, which gave dynamical masses of \qty{2.063+-0.023}{\solarmass} and \qty{1.018+-0.011}{\solarmass} for A and B, respectively.

To find the total age of Sirius B, \citet{bond_sirius_2017} used isochrones to constrain the cooling age (\qty{126}{\mega\year}), first. Applying the initial-final mass relation (IFMR) of white dwarfs \citep{cummings_two_2016} the estimated progenitor mass of Sirius B is between \qtyrange{5}{5.6}{\solarmass}, which, when combined with stellar evolution codes, yielded total system ages between \qtyrange{226}{228}{\mega\year} with an uncertainty of about \qty{+-10}{\mega\year} \citep{bond_sirius_2017}.

A companion around Sirius B would be affected by the orbit of Sirius A, and this constrained three-body system has been studied numerically \citep{holman_long-term_1999}. \citet{bond_sirius_2017} calculate the longest period stable companion around Sirius B is \qty{1.79}{\year}, which corresponds to a \qty{1.5}{\au} circular orbit.

@@

In this work we report direct images of Sirius B with Keck/NIRC2. \autoref{sec:obs} describes our target and observing strategy, \autoref{sec:analysis} describes the processing and analysis techniques used, \autoref{sec:results} describes our results and we conclude with \autoref{sec:conclusion}.

@@

\begin{deluxetable*}{ccccccccc}
    \tabletypesize{\small}
    \tablecaption{Observing parameters for the three epochs of data. All observations were carried out using the NIRC2 Lp-band filter. Observation time is based on the frames that were selected for processing. Seeing values are measured at \qty{0.5}{\micro\meter} using a differential image motion monitor (DIMM) and averaged over the observing session. Seeing values, temperature, and water vapor measurements are all provided by the Maunakea weather center forecast archive.
    \label{tbl:obs}}
    \tablehead{
        \colhead{\makecell{Date\\observed}} &
        \colhead{\makecell{Sirius B\\offset} (\unit{\arcsecond})} &
        \colhead{\makecell{Sirius B\\PA} (\unit{\degree})} &
        \colhead{\makecell{Obs.\\time} (hr)} &
        \colhead{\makecell{FOV\\rotation} (\unit{\degree})} &
        \colhead{FWHM (\unit{\milliarcsecond})} &
        \colhead{Seeing (\unit{\arcsecond})} &
        \colhead{Temp (\unit{\celsius})} &
        \colhead{PWV (\unit{\milli\meter})}
    }
    \startdata
    2020-02-04 & 11.20 & 67.90 & 1.44 & 60.1 & 79.9 & 0.936 & 0.0 & 0.7 \\
    2020-11-21 & 11.27 & 66.42 & 2.91 & 91.4 & 76.4 & 0.871 & 0.8 & 3.5 \\
    2020-11-28 & 11.27 & 66.38 & 2.44 & 80.4 & 82.2 & 1.23 & -1.5 & 3.0 \\
    \enddata
\end{deluxetable*}

@@

Despite Sirius B being the brightest white dwarf in the sky, it is still 10 magnitudes fainter than Sirius A, making it a technically challenging target, especially on ground-based telescopes. We directly imaged Sirius B with Keck/NIRC2 in Lp-band (\qty{3.776}{\micro\meter}) using the narrow camera (\qty{10}{\milliarcsecond\per px}; \ang{;;2.5}$\times$\ang{;;2.5}) across three epochs in 2020 (\autoref{tbl:obs}). Our first attempt to observe Sirius B failed due to the strong scattered light from Sirius A. The adaptive optics (AO) calibration failed when the scattered light from Sirius A would sweep into the FOV of the wavefront sensor (WFS). Similarly, trying to deploy the focal-plane vortex coronagraph \citep{serabyn_w_2017} failed when the coronagraphic pointing control algorithm, \texttt{QACITS} \citep{huby_w_2017}, performed erratically in the presence of the scattered light. \citet[\S2]{vigan_high-contrast_2015} reported similar issues in their attempts to image Sirius B coronagraphically. In order to overcome these obstacles, we decided to try using Sirius A as the AO guide star and off-axis guiding to Sirius B.

The Keck AO system \citep{wizinowich_performance_2000} was saturated by Sirius A, so we attenuated the flux using a narrow laser-line filter to avoid saturating the WFS. While still bright (appearing like a $\sim$5 magnitude star on the WFS), this was enough attenuation to close the AO loop. From here, we slewed off-axis using the separations and position angles calculated in \autoref{tbl:obs} from the orbital parameters in \citet{bond_sirius_2017}. In this mode, we noticed higher than usual drift in the focal plane, requiring manually recentering the target every 5 or 10 minutes. We tried deploying the vortex coronagraph along with \texttt{QACITS}, but the diffracted light from Sirius A made \texttt{QACITS} unstable, so we decided to forego any coronagraphy for the remaining observations.

@@

\begin{figure*}
    \centering
    \epsscale{1}
    \plotone{contrast_curves}
    \caption{The contrast curves for the best performing algorithm from each epoch. The solid lines are the Gaussian 5$\sigma$ contrast curves and the dashed lines are the Student-t corrected curves. The mean uncertainty from the total system age ($\sim$\qty{10}{\mega\year}) is shown by a lone error bar. In addition, the expected upper limit for orbital separation of a stable orbit of \qty{1.5}{\au} is plotted as a vertical dashed line. The companion mass values are interpolated from the AMES-Cond grid. The lower mass limit (upper magnitude limit) of these models is plotted in a light-gray horizontal dashed line. The annular PCA curve is cut off because the contrast in the innermost annulus was greater than 1.}
    \label{fig:contrast}
\end{figure*}

@@

The raw images from NIRC2 required pre-processing before analyzing them for companions. For each epoch we applied a flat correction using calibration frames captured during observing. We also removed bad pixels using a combination of \texttt{L.A.Cosmic} \citep{van_dokkum_cosmic-ray_2001} and an adaptive sigma-clipping algorithm. We removed sky background using a high-pass median filter with a box size of 31 pixels. For both the November epochs we tried exploiting the large focal plane drifts by dithering between two positions in order to simplify background subtraction, but this ended up performing worse than the high-pass filter. At this point frames were manually selected to remove bad frames, especially those with diffraction spikes from Sirius A within a few hundred pixels, like in \autoref{fig:spike}. Then, each good frame was co-registered to sub-pixel accuracy using the algorithm presented in \citet{guizar-sicairos_efficient_2008}, followed by fitting each frame with a Gaussian PSF to reach floating-point centroid accuracy.

The co-registered frames are then shifted to the center of the FOV. Lastly the frames were cropped to the inner 200 pixels and stacked into data cubes for each epoch. With a pixel scale of \qty{10}{\milliarcsecond} the crop corresponds to a maximum separation of \ang{;;1} or a projected separation of \qty{2.7}{\au}. We also measure the parallactic angle of each frame, including corrections for distortion effects following \citet{yelda_improving_2010}. For each epoch, we measure the full-width at half-maximum (FWHM) of the stellar PSF for use in post-processing by fitting a bivariate Gaussian model to the median frame from each data cube (\autoref{fig:psf}). All of the pre-processing code is available in Jupyter notebooks in a GitHub repository (\href{https://github.com/mileslucas/sirius-b}{https://github.com/mileslucas/sirius-b}) and the pre-processed data cubes and parallactic angles are available on Zenodo:\dataset[10.5281/zenodo.5115225]{\doi{10.5281/zenodo.5115225}}.

@@

We used three metrics for determining the performance of each algorithm, the signal-to-noise ratio (S/N) significance map, the standardized trajectory intensity mean map \citep[STIM map;][]{pairet_stim_2019}, and the contrast curve. The significance and STIM maps assign a likelihood to each pixel for the presence of a companion using different assumptions of the residual statistics. The contrast curve determines the sensitivity of a 5$\sigma$ statistical detection through repeated injection and retrieval of planetary signal as processed by one of the ADI algorithms above. We calculate both the Gaussian contrast and the Student-t corrected contrast, which accounts for the small-sample statistics in each annulus \citep{mawet_fundamental_2014}. The collapsed residual frames along with the above metrics for each algorithm for each epoch are in \autoref{sec:adi-results}.

A common problem when using subspace-driven post-processing algorithms like PCA, NMF, or GreeDS is choosing the size of the subspace (i.e., the number of components). For PCA, NMF, and GreeDS algorithms, we created residual cubes for increasing number of components, from 1 to 10. We chose 10 for the max number of components because we saw a dramatic decline in contrast sensitivity after the first few components (\autoref{fig:pca-contrast-curves}). In our analysis we employed the STIM largest intensity mask map \citep[SLIM map;][]{pairet_signal_2020} as an ensemble statistic. The SLIM map calculates the average STIM map from many residual cubes along with the average mask of the $N$ most intense pixels in each STIM map. A real companion ought to be present in many different residual cubes from the same dataset, so this ensemble approach can give us a probability map without predetermining the number of components. The collapsed residual frames, average STIM map, SLIM map, and contrast curves for each epoch for each of the above algorithms are in \autoref{sec:adi-results}.

@@

We determined the best-performing algorithms for each epoch using the contrast curves described in \autoref{sec:analysis}. For the first two epochs full-frame median subtraction had the best contrast at almost all separations. For the last epoch annular PCA subtraction with 2 principal components and a rotation threshold of 0.5 FWHM produced the best contrast at close separations (\qtyrange{0.2}{0.4}{\arcsecond}) and had similar performance to other algorithms beyond \ang{;;0.4}. The innermost annulus from this algorithm has invalid contrast ($>$1) and is not shown. The collapsed residual frames from each epoch are shown in \autoref{fig:residuals}, along with the Gaussian significance maps (\autoref{fig:sig}) and STIM maps (\autoref{fig:stim}).

The reduced images do not show consistent or significant evidence for a substellar companion. The STIM probability maps for the 2020-11-21 and 2020-11-28 epochs suggest evidence for some blobs $\sim$\qty{0.3}{\au} (\ang{;;0.13}; 1.6 FWHM) from the center. The February epoch also shows a blob at a similar separation in the reduced image which does not appear in the STIM map. The lack of statistical evidence in the February epoch and the significance maps as well as the proximity to the central star both reduce the probability of these blobs being true companions. Nonetheless, we estimated astrometry for blobs from each epoch (\autoref{tbl:astrometry}) and tried fitting Keplerian orbits using the ``Orbits for the Impatient'' algorithm \citep[OFTI;][]{blunt_orbits_2017}. We generated $10^4$ orbits, none of which managed to contain the points from each epoch (\autoref{sec:orbits}). We take this as direct evidence against the blobs being substellar companions of any kind.

>>>>

\documentclass[twocolumn]{aastex631}
\usepackage{cleveref}
\let\tablenum\relax
\usepackage{siunitx}
\usepackage{makecell}

% workaround for cleveref+AASTeX
% https://github.com/AASJournals/AASTeX60/issues/69#issuecomment-685852891
\makeatletter
\usepackage{etoolbox}
\patchcmd\H@refstepcounter{\protected@edef}{\protected@xdef}{}{}
\makeatother

% custom SI units
\DeclareSIUnit\parsec{pc}
\DeclareSIUnit\au{AU}
\DeclareSIUnit\jupitermass{M_J}
\DeclareSIUnit\milliarcsecond{mas}
\DeclareSIUnit\solarmass{M_\odot}
\DeclareSIUnit\year{yr}
\DeclareSIUnit\pixel{px}
\sisetup{
    list-units=single,
    separate-uncertainty
}

@@

\begin{abstract}
We present deep imaging of Sirius B, the closest and brightest white dwarf. We use Keck/NIRC2 in Lp-band (\qty{3.776}{\micro\meter}) across three epochs in 2020 using the technique of angular differential imaging. We reach sub-Jupiter sensitivities and sub-AU separations, reaching \qty{3.5}{\jupitermass} at \qty{0.25}{\au} down to a sensitivity of \qty{0.6}{\jupitermass} at \qty{1.5}{\au}. The uncertainty in mass sensitivity is below \qty{0.1}{\jupitermass} due to the precisely known Sirius B system age. Consistent with previous results, we do not detect any companions around Sirius B.
\end{abstract}

@@

High-contrast imaging (HCI) is a powerful technique for discovering and characterizing exoplanets. Probing the architecture, formation, and atmospheres of planets directly is necessary for advancing substellar companion formation and evolution theory. Imaging a planet is challenging, however. The typical astrophysical flux ratios (contrast) for a Sun-Jupiter analog in the near-infrared (NIR) are $\sim$$10^{\text{-}8}$, and for a Sun-Earth system are $\sim$$10^{\text{-}10}$ \citep{traub_direct_2010}. Thermal emission from exoplanets peaks in the infrared and is well into the Rayleigh-Jeans limit of the star, decreasing contrast compared to the visible or ultraviolet. In addition, the close angular separations of planets make it difficult to detect them over the diffraction pattern of their host and other noise sources.

Typical targets for imaging are nearby young stars; the proximity means the same angular separation probes a smaller physical separation, allowing for studies closer to solar-system scales. Younger exoplanets are hotter and therefore brighter, reducing the flux ratio between the planet and its host star. Another way to reduce contrast would be to image a faint star with bright companions, such as a young white dwarf (WD) star. When intermediate-mass stars (\qtyrange{1}{8}{\solarmass}) eventually exit the main-sequence (MS) they expand up to $\sim$100s of stellar radii during the red giant branch (RGB) phase. This expansion engulfs planets within the star's Roche limit. If the planet survives this expansion, its orbital separation will gradually increase as the star loses mass during the asymptotic giant branch (AGB) phase. Finally, once the star becomes a white dwarf it begins cooling, which will reduce its luminosity by 3 to 4 orders of magnitude compared to its MS progenitor. The faintness, lack of spectral features, and expanded orbits make white dwarfs exceptionally challenging for the transit photometry and radial velocity methods. However, the reduced contrast due to the faintness of the star, and the expanded separation of planets both benefit imaging.

@@

There is limited knowledge of planetary systems around evolved stars. \citet{burleigh_imaging_2002,veras_post-main-sequence_2016} suggest exoplanets on initially wide (\textgreater\qty{5}{\au}) orbits around intermediate-mass stars will survive expansion during the RGB phase. Planets very close to the Roche limit of the expanding red giant can still be shredded by tidal forces despite escaping engulfment, although the tidal forces are negligible on planets that are far from the Roche limit (and readily escape engulfment) \citep{nordhaus_orbits_2013}. During the AGB phase, stellar mass loss adiabatically expands the orbit of the planet by a maximum factor of $M_{*,\mathrm{MS}}/M_{*,\mathrm{WD}}$ \citep{jeans_cosmogonic_1924}. In addition to \textit{first-generation} planets, there are potential methods for \textit{second-generation} planets to form after the violent RGB and AGB phases \citep{perets_second_2010}.

The first search for substellar companions around white dwarfs was conducted by \citet{probst_infrared_1983}. They searched for infrared (IR) excess in their spectral energy distributions (SED) using broadband photometry. The IR excess would be indicative of thermal emission of faint material around the star, such as a circumstellar disk or giant planet. They studied $\sim$100 white dwarfs but found no companions. The same method was applied by \citet{zuckerman_excess_1987}, who found excess IR emission around white dwarf G29-38. The emission was determined to be from a dust disk \citep{telesco_observations_1990}, which was associated with accretion onto the white dwarf's surface, polluting the stellar atmosphere \citep{koester_metals_1997}. It is estimated that while a third of white dwarfs are polluted, only a few percent have IR excesses \citep{bonsor_pollution_2015}. \citet{skemer_sirius_2011} suggested stellar radiation from stars hotter than \qty{15,000}{\kelvin} would sublimate dust therefore suppressing any IR excess.

Imaging is well-suited for studying evolved planetary systems, yet to date, only one planet has been imaged around a white dwarf. \citet{luhman_discovery_2011} discovered a companion around WD 0806-661B with a mass of \qty{7}{\jupitermass} on a wide \qty{2500}{\au} orbit using Spitzer. The ``Degenerate Objects around Degenerate Objects'' survey \citep[DODO;][]{hogan_dodo_2009} observed 29 white dwarfs with Gemini/NIRI and VLT/ISAAC, reaching an average upper limit around $\sim$\qty{8}{\jupitermass} beyond \qty{35}{\au}. The young white dwarf GD 50 was observed using the extreme AO instrument SPHERE at the VLT \citep{xu_extreme-ao_2015}, reaching sensitivity limits of \qty{4}{\jupitermass} at \qty{6.2}{\au}.

@@

A particularly fascinating such target is Sirius B, the closest and brightest white dwarf. The Sirius system is the 7th closest to the sun at \qty{2.6}{\parsec}, consisting of Sirius A, a -1.35 magnitude A1Vm star known for being the brightest, and Sirius B, a DA2 white dwarf with a \qty{50}{\year} orbit \citep{bond_sirius_2017,gaia_collaboration_gaia_2018}. As mentioned previously, the proximity and faintness of Sirius B (compared to a MS star) make it compelling for imaging. Additionally, the young system age ($\sim$\qty{225}{\mega\year}) means any giant planets would still retain much of their latent formation heat, increasing their luminosity in the IR \citep{fortney_giant_2010}.

A common problem in direct imaging is determining planetary masses from photometry using planetary atmosphere evolution grids. This method of interpolation depends on the stellar age, the stellar photometry, and the distance to the system. Determining stellar ages is quite difficult and sometimes impossible, and is often the largest source of uncertainty when interpolating planetary masses. Age uncertainty exponentially affects young systems (\textless\qty{1}{\giga\year}) due to the rapid cooling of giant planets. Sirius B avoids these pitfalls due to its unusually high age precision, primarily derived from accurate dynamical masses.

The dynamical masses are determined through astrometric studies; the first study of Sirius was performed by \citet{bessel_variations_1844}, who recognized wobbles in the proper motion of Sirius A caused by a ``dark satellite''. This dark satellite was visually confirmed in \citet{bond_companion_1862} as Sirius B. \citet{adams_spectrum_1915} took the first spectral measurements of Sirius B and found it to be similar to a MS early A-type star, despite its faintness, which we now know to be typical of white dwarf spectra.

Initial astrometric measurements suggested a 50 year orbital period for Sirius B \citep{auwers_orbit_1864}. \citet{van_den_bos_orbit_1960,gatewood_study_1978} were the first to estimate dynamical masses using compilations of photographic plates from Lick and Yerkes observatories. \citet{bond_sirius_2017} greatly refined the orbital solution using a compilation of historical data and Hubble Space Telescope (HST) data, which gave dynamical masses of \qty{2.063+-0.023}{\solarmass} and \qty{1.018+-0.011}{\solarmass} for A and B, respectively. A companion around Sirius B would be affected by the orbit of Sirius A, and this constrained three-body system has been studied numerically \citep{holman_long-term_1999}. \citet{bond_sirius_2017} calculate the longest period stable companion around Sirius B is \qty{1.79}{\year}, which corresponds to a \qty{1.5}{\au} circular orbit.

To find the total age of Sirius B, \citet{bond_sirius_2017} used isochrones to constrain the cooling age (\qty{126}{\mega\year}), first. Applying the initial-final mass relation (IFMR) of white dwarfs \citep{cummings_two_2016} the estimated progenitor mass of Sirius B is between \qtyrange{5}{5.6}{\solarmass}, which, when combined with stellar evolution codes, yielded total system ages between \qtyrange{226}{228}{\mega\year} with an uncertainty of about \qty{+-10}{\mega\year} \citep{bond_sirius_2017}. An age uncertainty of $\sim$1\% is execeptional compared to the $\sim$10\% or worse of stars found in young moving groups, not to mention many stars cannot be precisely dated at all.

@@

In this work we report direct images of Sirius B with Keck/NIRC2. The rest of the manuscript is organized as follows: \cref{sec:obs} describes our target and observing strategy, \cref{sec:analysis} describes the processing and analysis techniques used, and \cref{sec:results} describes our results.

@@

\begin{deluxetable*}{ccccccccc}
    \tabletypesize{\small}
    \tablecaption{Observing parameters for the three epochs of data. All observations were carried out using the NIRC2 narrow camera (\qty{10}{\milliarcsecond\per\pixel}; \ang{;;2.5}$\times$\ang{;;2.5}) in Lp-band (\qty{3.776}{\micro\meter}). Observation time is based on the frames that were selected for processing. Seeing values were measured at \qty{0.5}{\micro\meter} using a differential image motion monitor (DIMM) and averaged over the observing session. Seeing values, temperature, and water vapor measurements were all taken from the Maunakea weather center forecast archive.
    \label{tbl:obs}}
    \tablehead{
        \colhead{\makecell{Date\\observed}} &
        \colhead{\makecell{Sirius B\\offset} (\unit{\arcsecond})} &
        \colhead{\makecell{Sirius B\\PA} (\unit{\degree})} &
        \colhead{\makecell{Obs.\\time} (hr)} &
        \colhead{\makecell{FOV\\rotation} (\unit{\degree})} &
        \colhead{FWHM (\unit{\milliarcsecond})} &
        \colhead{Seeing (\unit{\arcsecond})} &
        \colhead{Temp (\unit{\celsius})} &
        \colhead{PWV (\unit{\milli\meter})}
    }
    \startdata
    2020-02-04 & 11.20 & 67.90 & 1.44 & 60.1 & 79.9 & 0.936 & 0.0 & 0.7 \\
    2020-11-21 & 11.27 & 66.42 & 2.91 & 91.4 & 76.4 & 0.871 & 0.8 & 3.5 \\
    2020-11-28 & 11.27 & 66.38 & 2.44 & 80.4 & 82.2 & 1.23 & -1.5 & 3.0 \\
    \enddata
\end{deluxetable*}

@@

Despite Sirius B being the brightest white dwarf in the sky, it is still 10 magnitudes fainter than Sirius A, making it a technically challenging target, especially on ground-based telescopes. We directly imaged Sirius B with Keck/NIRC2 in Lp-band (\qty{3.776}{\micro\meter}) using the narrow camera (\qty{10}{\milliarcsecond\per\pixel}; \ang{;;2.5}$\times$\ang{;;2.5}) across three epochs in 2020 (\cref{tbl:obs}). Our first attempt to observe Sirius B failed due to the strong scattered light from Sirius A. The adaptive optics (AO) calibration failed when the scattered light from Sirius A would sweep into the FOV of the wavefront sensor (WFS). Similarly, trying to deploy the focal-plane vortex coronagraph \citep{serabyn_w_2017} failed when the coronagraphic pointing control algorithm, \texttt{QACITS} \citep{huby_w_2017}, performed erratically in the presence of the scattered light. We did not try coronagraphy for the remaining observations. \citet[\S2]{vigan_high-contrast_2015} reported similar issues in their attempts to image Sirius B coronagraphically using VLT/SPHERE. In order to overcome these obstacles, we decided to try using Sirius A as the AO guide star and off-axis guiding to Sirius B.

The Keck AO system \citep{wizinowich_performance_2000} was saturated by Sirius A, so we attenuated the flux using a narrow laser-line filter in the WFS. While still bright (appearing like a $\sim$5 magnitude star on the WFS), this was enough attenuation to close the AO loop. From here, we slewed off-axis using the separations and position angles calculated in \cref{tbl:obs} from the orbital parameters in \citet{bond_sirius_2017}. In this mode, we noticed higher than usual drift in the focal plane, requiring manually recentering the target every 5 or 10 minutes.

@@

\begin{figure*}
    \centering
    \epsscale{1}
    \plotone{contrast_curves}
    \caption{The contrast curves for the best performing algorithm from each epoch. The solid lines are the Gaussian 5$\sigma$ contrast curves and the dashed lines are the Student-t corrected curves. The mean uncertainty from the total system age ($\sim$\qty{10}{\mega\year}) is shown by a lone error bar. In addition, the expected upper limit for orbital separation of a stable orbit of \qty{1.5}{\au} is plotted as a vertical dashed line. The companion mass values are interpolated from the AMES-Cond grid. The lower mass limit (upper magnitude limit) of these models is plotted in a light-gray horizontal dashed line. The annular PCA curve is cut off because the innermost annulus had greater than 1 contrast, which is invalid.}
    \label{fig:contrast}
\end{figure*}

@@

The raw images from NIRC2 required pre-processing before analyzing them for companions. For each epoch we applied a flat correction using calibration frames captured during observing. We also removed bad pixels using a combination of \texttt{L.A.Cosmic} \citep{van_dokkum_cosmic-ray_2001} and an adaptive sigma-clipping algorithm. We removed sky background using a high-pass median filter with a box size of 31 pixels. For both the November epochs we tried exploiting the large focal plane drifts by dithering between two positions in order to simplify background subtraction, but this ended up performing worse than the high-pass filter. At this point frames were manually selected to remove bad frames, especially those with diffraction spikes from Sirius A within a few hundred pixels, like in \cref{fig:spike}. Then, each good frame was co-registered to sub-pixel accuracy using the algorithm presented in \citet{guizar-sicairos_efficient_2008}, followed by fitting each frame with a Gaussian PSF to further increase centroid accuracy.

The co-registered frames were shifted to the center of the FOV and cropped to the inner 200 pixels. With a pixel scale of \qty{10}{\milliarcsecond} the crop corresponds to a maximum separation of \ang{;;1} or a projected separation of \qty{2.7}{\au}. All the frames were stacked into data cubes for each epoch. We also measure the parallactic angle of each frame, including corrections for distortion effects following \citet{yelda_improving_2010}. For each epoch, we measure the full-width at half-maximum (FWHM) of the stellar PSF for use in post-processing by fitting a bivariate Gaussian model to the median frame from each data cube (\cref{fig:psf}). All of the pre-processing code is available in Jupyter notebooks in a GitHub repository (\href{https://github.com/mileslucas/sirius-b}{https://github.com/mileslucas/sirius-b}) and the pre-processed data cubes and parallactic angles are available on Zenodo:\dataset[10.5281/zenodo.5115225]{\doi{10.5281/zenodo.5115225}}.

@@

We used three metrics to determine the performance of each algorithm, the signal-to-noise ratio (S/N) significance map, the standardized trajectory intensity mean map \citep[STIM map;][]{pairet_stim_2019}, and the contrast curve. The significance and STIM maps assign a likelihood to each pixel for the presence of a companion using different assumptions of the residual statistics. The contrast curve determines the sensitivity of a 5$\sigma$ statistical detection through repeated injection and retrieval of planetary signal as processed by one of the ADI algorithms above. We calculate both the Gaussian contrast and the Student-t corrected contrast, which accounts for the small-sample statistics in each annulus \citep{mawet_fundamental_2014}. The collapsed residual frames along with the above metrics for each algorithm for each epoch are in \cref{sec:adi-results}.

A common problem when using subspace-driven post-processing algorithms like PCA, NMF, or GreeDS is choosing the size of the subspace (i.e., the number of components). For PCA, NMF, and GreeDS algorithms, we created residual cubes for increasing number of components, from 1 to 10. We chose 10 for the maximum number of components because we saw a dramatic decline in contrast sensitivity after the first few components (\cref{fig:pca-contrast-curves}). In our analysis we employed the STIM largest intensity mask map \citep[SLIM map;][]{pairet_signal_2020} as an ensemble statistic. The SLIM map calculates the average STIM map from many residual cubes along with the average mask of the $N$ most intense pixels in each STIM map. A real companion ought to be present in many different residual cubes from the same dataset, so this ensemble approach can give us a probability map without predetermining the number of components. The collapsed residual frames, average STIM map, SLIM map, and contrast curves for each epoch for each of the above algorithms are in \cref{sec:adi-results}.

@@

We determined the best-performing algorithms for each epoch using the contrast curves described in \cref{sec:analysis}. For the first two epochs full-frame median subtraction had the best contrast at almost all separations. For the last epoch annular PCA subtraction with 2 principal components and a rotation threshold of 0.5 FWHM produced the best contrast at close separations (\qtyrange{0.2}{0.4}{\arcsecond}) and had similar performance to other algorithms beyond \ang{;;0.4}. The innermost annulus from this algorithm has invalid contrast (\textgreater1) and is not shown. The collapsed residual frames from each epoch are shown in \cref{fig:residuals}, along with the Gaussian significance maps (\cref{fig:sig}) and STIM maps (\cref{fig:stim}).

The reduced images do not show consistent or significant evidence for a substellar companion. The STIM probability maps for the 2020-11-21 and 2020-11-28 epochs suggest evidence for some blobs $\sim$\qty{0.3}{\au} (\ang{;;0.13}; 1.6 FWHM) from the center. The February epoch also shows a blob at a similar separation in the reduced image which does not appear in the STIM map. The lack of statistical evidence in the February epoch and the significance maps as well as the proximity to the central star both reduce the probability of these blobs being true companions. Nonetheless, we estimated astrometry for blobs from each epoch (\cref{tbl:astrometry}) and tried fitting Keplerian orbits using the ``Orbits for the Impatient'' algorithm \citep[OFTI;][]{blunt_orbits_2017}. We generated $10^4$ orbits, none of which contained the points from each epoch (\cref{sec:orbits}). We take this as direct evidence against the blobs being substellar companions of any kind.
